---
title: "My Project"
format: html
editor: visual
---

## Uganda Waterpoints

This dataset provides comprehensive information on **water points across Uganda**. This dataset is a free dataset shared through [The Humanitatian Data Exchange](https://data.humdata.org/dataset/wpdx_uga "Uganda Water points Data"). Each entry represents an individual water source, such as a well or a rainwater harvesting system, with associated details covering location, type, functionality, management, and usage.

Key attributes include:

-    **Geographical information:** Latitude, longitude, and administrative divisions (region, district, sub-county).

-   **Water source and technology:** The type of water source (e.g., protected well, rainwater harvesting) and the technology used.

-   **Operational status:** Whether the water point is functional or non-functional.

-   **Management and governance:** How the water point is managed (community-managed, fee-based, etc.) and the organization responsible.

-   **Installation and maintenance history:** Year of installation, rehabilitation records, and organizations involved.

-   **Population served:** The number of people served within the vicinity of the water point.

-   **Water quality indicators:** Includes parameters such as fecal contamination presence and subjective quality assessments.

-   **Proximity metrics:** Distances to urban centers, primary, secondary, and tertiary water sources.

-   **Predictive assessments:** Predicted functionality over time, providing insights into future reliability.

This dataset serves as a detailed resource for monitoring water access and quality in Uganda, supporting analysis of water point functionality, population coverage, management effectiveness, and planning for rehabilitation or new installations.

## Import

To load the data, you should first load the required libraries to use:

```{r}
library(tidyverse)
```

Now, load your dataset like this

```{r}
#| echo: false
read_csv(here::here("data/raw/wpdx_enhanced.csv"))
```
